\section{Abstract}

This document describes Monte Carlo simulation of intermediate energy nuclear reactions of hadronic cascade models in {\sc Geant4}.
Two spesific codes are described, namely HETC and INUCL. 
Models treated are actual intra-nuclear cascade models, fission models, pre-equilibrium, and evaporation models.
We represent implementation details and results achived from simulations and make comparisons with measurements.
This document also works as physics manual, status report and working document for the project.


We present Geant4 implementation of Bertini intranuclear cascade model based on High Energy Transport Code (HETC)

We describe here project to implement High Energy Transport Code (HETC) software into Geant4. 
We give on overview of the hadronic Monte Carlo models used and present first results from evaporation part of HETC. 
We also outline project goals and future development.

HIP has requested (Spring 2001) to join Geant4 (G4) collaboration. 
HIP Software and Physics Group will contibute Geant4 development with annual work of two man years. 
This year 2001, our commitment is fulfilled with efforts made by Aatos
Heikkinen, Nikita Stepanov, Laura Tillikainen and Tapio Lampen.

One part of our agreement is that HIP provides C++ translation of Fortran based High Energy Transport Code (HETC) into G4 before 2002.

Our work consists of re-engineering, implementing, validating and documenting the HETC into new module (called HETC++) in G4 hadronic models framework.



HETC/INUCL is a simulation package and a model of intermediate energy nuclear
reactions.
INUCL consists of cascade, fission models, pre-equilibrium, and evaporation
This manual describes the models used in INUCL++. 
This documet works as an manual, status report and working document for INUCL project. In accordance with this, we represent results acieved with INUCL. Comparison with measurement is made. 
We also descripe staus of Geant4 installation on INUCL software \cite{titarenko99a}.

